---
title: "Descriptives and tests of coarse grained samples Geology, GEO-112, Geography Degree, UASD"
output:
  pdf_document
  # html_document
header-includes:
  \usepackage{pdflscape}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r, tidy=F}
library(readODS)
library(sp)
library(tidyverse)
library(plotKML)
```

## Set the working directory, read the data, prepare the data

```{r, tidy=F}
#Working directory
wd <- '/home/jr/Documentos/clases_UASD/sem201802/datos_campo/'
setwd(wd)

#Spatial object
fcoords <- list.files(wd, '*coord.ods')
coords <- read_ods(
  fcoords,
  col_types = c('text', 'text', 'numeric', 'numeric', 'text')
)
str(coords)
coords[,c(3,4)] <- sapply(coords[,c(3,4)], as.numeric)
str(coords)

#Samples
fsamples <- list.files(wd, '*samples.ods')
samples <- read_ods(fsamples, range = 'A1:G1140')
str(samples)
samples[,c(5,6,7)] <- sapply(samples[,c(5,6,7)], as.numeric)
str(samples)
# Two letters abbreviation for responsible name
samples$responsable.abbv <- gsub(
  '\\b(\\pL)\\pL{2,}|.','\\U\\1',
  iconv(samples$responsable, to='ASCII//TRANSLIT'),
  perl = TRUE)
# Responsible.sample field (e.g. ET.M1)
samples$resp.m <- paste(
  samples$responsable.abbv,
  samples$`nombre de muestra`, sep = '.')
```

## Descriptive statistics and inferences

### Descriptive statistics

```{r, tidy=F}
sapply(
  unique(samples$responsable),
  function(x)
    psych::describe(samples[samples$responsable==x,c(5,6,7)]),
  simplify = F
)
sapply(
  unique(samples$responsable),
  function(x)
    sapply(
      unique(samples[samples$responsable==x,'nombre de muestra']),
      function(y)
        psych::describe(
          samples[samples$responsable==x&samples$`nombre de muestra`==y,c(5,6,7)]
          ),
      simplify = F
    )
)
```

### Inferences. Combinations of responsibles-samples

```{r, tidy=FALSE}
comb <- as.data.frame(
  gtools::combinations(
    length(unique(samples$resp.m)),
    r= 2,
    unique(samples$resp.m)
  )
)
head(comb)
```

### Inferences. t-Tests between responsibles-samples (all against all)
The smaller the p-value the better the result

```{r, tidy=FALSE}
rownames(comb) <- paste(comb[,1], comb[,2], sep = ' - ')
ttests <- t(
  sapply(
    rownames(comb),
    function(x)
      sapply(c('a','b','c'),
             function(y)
               t.test(
                 samples[samples$resp.m==comb[x,1], y],
                 samples[samples$resp.m==comb[x,2], y]
                 )$p.value,
             simplify = T,
             USE.NAMES = T
             )
    )
)
ttests
```

### Inferences. Number and relative frequencies of statistically significant differences between samples by responsibles (all against all)
- The higher the TRUE relative frequencies the better the result

```{r, tidy=F}
respnames <- unique(
  samples$responsable.abbv)[
    nchar(unique(samples$responsable.abbv))==2
    ]#This excludes Ocoa samples
ttests.nrlf <- map(
  respnames,
  ~ttests %>%
    as.data.frame() %>%
    rownames_to_column('m1 - m2') %>%
    mutate_if(is.numeric, funs(.<=0.05)) %>%
    filter(grepl(.x, `m1 - m2`)) %>% 
    filter(!grepl('Ocoa', `m1 - m2`)) %>% 
    gather(variable, value, -`m1 - m2`) %>% 
    dplyr::select(variable, value) %>% 
    group_by(variable, value) %>% 
    dplyr::summarise(freq=n()) %>%
    dplyr::add_tally(freq) %>% 
    mutate(proportion=round(freq/n*100,2))
) %>% set_names(respnames)
ttests.nrlf
```

### Inferences. Number and relative frequencies of statistically significant differences pairwise by responsible
- The higher the TRUE relative frequencies the better the result

```{r, tidy=F}
respnamesc <- as.data.frame(gtools::combinations(length(respnames), r=2, respnames))
rownames(respnamesc) <- paste(
  respnamesc$V1,
  '-',
  respnamesc$V2,
  '=',
  respnamesc$V2,
  '-',
  respnamesc$V1
)
ttests.nrlf.p <- map(
  rownames(respnamesc),
  ~ttests %>%
    as.data.frame() %>%
    rownames_to_column('m1 - m2') %>%
    mutate_if(is.numeric, funs(.<=0.05)) %>%
    filter(
      grepl(
        paste0(
          '^',
          respnamesc[.x,1],
          '\\.M. - ',
          respnamesc[.x,2],
          '.*$|^',
          respnamesc[.x,2],
          '\\.M. - ',
          respnamesc[.x,1],
          '.*$'),
        `m1 - m2`)) %>% 
    gather(variable, value, -`m1 - m2`) %>% 
    dplyr::select(variable, value) %>% 
    group_by(variable, value) %>% 
    dplyr::summarise(freq=n()) %>%
    dplyr::add_tally(freq) %>% 
    mutate(proportion=round(freq/n*100,2))
) %>% setNames(rownames(respnamesc))
ttests.nrlf.p
```

### Inferences. Statistically significant differences between samples of the same person
- The higher the TRUE relative frequencies the better the result

```{r, tidy=FALSE}
ttests.nrlf.s <- map(
  respnames,
  ~ttests %>%
    as.data.frame() %>%
    rownames_to_column('m1 - m2') %>%
    mutate_if(is.numeric, funs(.<=0.05)) %>%
    filter(grepl(
      paste0(
          '^',
          .x,
          '\\.M. - ',
          .x,
          '.*$')
      , `m1 - m2`)) %>% 
    gather(variable, value, -`m1 - m2`) %>% 
    dplyr::select(variable, value) %>% 
    group_by(variable, value) %>% 
    dplyr::summarise(freq=n()) %>%
    dplyr::add_tally(freq) %>% 
    mutate(proportion=round(freq/n*100,2))
) %>% set_names(respnames)
ttests.nrlf.s
```

<!-- ### Inferences. Correlations by axis -->

<!-- ```{r, tidy=F} -->
<!-- # a-axis -->
<!-- cor.a <- map( -->
<!--   rownames(respnamesc), -->
<!--   ~cor.test( -->
<!--     sort(samples[samples$responsable.abbv==respnamesc[.x,1],'a']), -->
<!--     sort(samples[samples$responsable.abbv==respnamesc[.x,2],'a']) -->
<!--   ) -->
<!-- ) %>% setNames(rownames(respnamesc)) -->
<!-- cor.a -->
<!-- # Correlation b-axis -->
<!-- cor.b <- map( -->
<!--   rownames(respnamesc), -->
<!--   ~cor.test( -->
<!--     sort(samples[samples$responsable.abbv==respnamesc[.x,1],'b']), -->
<!--     sort(samples[samples$responsable.abbv==respnamesc[.x,2],'b']) -->
<!--   ) -->
<!-- ) %>% setNames(rownames(respnamesc)) -->
<!-- cor.b -->
<!-- # Correlation c-axis -->
<!-- cor.c <- map( -->
<!--   rownames(respnamesc), -->
<!--   ~cor.test( -->
<!--     sort(samples[samples$responsable.abbv==respnamesc[.x,1],'c']), -->
<!--     sort(samples[samples$responsable.abbv==respnamesc[.x,2],'c']) -->
<!--   ) -->
<!-- ) %>% setNames(rownames(respnamesc)) -->
<!-- cor.c -->
<!-- ``` -->

### Plots

```{r respon.dens, tidy=FALSE, echo=TRUE, results='hide', fig.width=11, fig.show='hide'}
# Density plots
respon.dens <- samples %>%
  filter(!grepl('Ocoa', `nombre de muestra`)) %>% 
  select(responsible=responsable, a, b, c) %>% 
  gather(axis, `value (in mm)`, -responsible) %>% 
  ggplot(aes(x=`value (in mm)`, colour = responsible)) +
  geom_line(alpha = 0.75, stat = 'density', size = 2, fill = 'gray95') +
  facet_grid(~axis, scales = 'free_x') +
  theme(
    text = element_text(size = 18),
    axis.text.x=element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5),
    panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_line(colour = "grey", linetype = "dashed", size = 0.25),
    strip.background = element_rect(colour = "black", fill = "black"),
    strip.text.x = element_text(colour = "white", face = "bold")
  )
respon.dens
```

\newpage
\newgeometry{margin=0.1in, top=0.1in, headheight=0.0in, footskip=0.2in, includehead, includefoot}
\blandscape
```{r respon.dens2, echo=FALSE, tidy=FALSE, out.width='0.95\\paperheight', fig.width=11, fig.height=8, fig.align='center'}
# Box-Plots of species abundances (and transformations) by species
respon.dens
```
\elandscape
\restoregeometry

&nbsp;

```{r respon.bp, tidy=FALSE, echo=TRUE, results='hide', fig.width=11, fig.show='hide'}
# Box-plots
respon.bp <- samples %>%
  filter(!grepl('Ocoa', `nombre de muestra`)) %>% 
  select(responsible=responsable, a, b, c) %>% 
  gather(axis, `value (in mm)`, -responsible) %>% 
  ggplot(aes(x=responsible, y=`value (in mm)`, fill = responsible)) +
  geom_boxplot(alpha = 0.3) +
  facet_grid(~axis, scales = 'free_x') +
  theme(
    text = element_text(size = 18),
    axis.text.x=element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5),
    panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_line(colour = "grey", linetype = "dashed", size = 0.25),
    strip.background = element_rect(colour = "black", fill = "black"),
    strip.text.x = element_text(colour = "white", face = "bold")
  )
respon.bp
```

\newpage
\newgeometry{margin=0.1in, top=0.1in, headheight=0.0in, footskip=0.2in, includehead, includefoot}
\blandscape
```{r respon.bp2, echo=FALSE, tidy=FALSE, out.width='0.95\\paperheight', fig.width=11, fig.height=8, fig.align='center'}
# Box-Plots of species abundances (and transformations) by species
respon.bp
```
\elandscape
\restoregeometry


### Maps

```{r, tidy=FALSE}
# coords.sp <- coords
# coordinates(coords.sp) <- ~x+y
# proj4string(coords.sp) <- CRS("+init=epsg:32619")
# plotKML(coords.sp, 'kml')
# spTransform(coords.sp, CRS("+init=epsg:4326"))
```
